##Load Data

```{r}
library(gtools)
library(ggplot2)
library(ggsci)
library(tidyverse)
library(gridExtra)
library(scales)
library(cluster)
library(stringr)
library(ggsignif)
library(forcats)
rm(list=ls())

setwd('C:/Users/cmorc/OneDrive - Harvard University/Projects/Metabolic Aging Atlas/C.-elegans-Mitochondrial-Aging/DR Analyses/')

data_mito <- read.csv('C:/Users/cmorc/OneDrive - Harvard University/Projects/Metabolic Aging Atlas/C.-elegans-Mitochondrial-Aging/FLIM_Mito_052623.csv')

data_mito_animal <- read.csv('C:/Users/cmorc/OneDrive - Harvard University/Projects/Metabolic Aging Atlas/C.-elegans-Mitochondrial-Aging/FLIM_Mito_Median_Per_Animal_052623.csv')

data_mito_animal_matrix <- read.csv('C:/Users/cmorc/OneDrive - Harvard University/Projects/Metabolic Aging Atlas/C.-elegans-Mitochondrial-Aging/FLIM_Mito_Median_Per_Animal_Animal_Matrix_052623.csv')

```

##Process Data for plots
```{r}
data_mito_animal$Condition <- str_c(data_mito_animal$Age," ", data_mito_animal$TissueFullName," ",data_mito_animal$Nutrients)

myagesoy <- c("1","11")
data_aldr_youngold <- as.data.frame(subset(data_mito_animal_matrix, Age%in%myagesoy))
data_aldr_youngold$Age <- as.character(data_aldr_youngold$Age)
data_aldr_youngold$Condition <- str_c("Day ",data_aldr_youngold$Age," ",data_aldr_youngold$Nutrients)

data_mito_animal$Imean <- ave(data_mito_animal$i, data_mito_animal$Condition)
data_mito_animal$a1mean <- ave(data_mito_animal$a1, data_mito_animal$Condition)
data_mito_animal$t1mean <- ave(data_mito_animal$t1, data_mito_animal$Condition)
data_mito_animal$t2mean <- ave(data_mito_animal$t2, data_mito_animal$Condition)
data_mito_animal$tmmean <- ave(data_mito_animal$tm, data_mito_animal$Condition)
myvars2 <- c("a1mean", "t2mean", "t1mean", "Imean", "tmmean", "Condition", "Age", "Nutrients", "Tissue", "TissueFullName")
data_averages <- unique(data_mito_animal[myvars2], by ="Imean")
```

##Plot bwm tm D1 vs 11, AL vs DR
```{r}
data_aldr_youngold_musNAremov <- data_aldr_youngold[data_aldr_youngold$t.bwm.tm != "#N/A", ] 
data_aldr_youngold_musNAremov$t.bwm.tm <- as.numeric(data_aldr_youngold_musNAremov$t.bwm.tm)

pdf(paste(Sys.Date(),"AL DR Y O bwm.pdf"))
ggplot(data=data_aldr_youngold_musNAremov, aes(Age, t.bwm.tm, fill=Nutrients))+
geom_violin()+
stat_summary(fun.data="mean_cl_boot", size=1, position = position_dodge(width = 0.9))+
ylab(bquote("B.A./NAD(P)H  " *tau[m]* " (ns)"))+
xlab("Age (Days)")+
theme_classic(base_size = 20, base_family = "")+
theme(aspect.ratio=1)+
theme(text=element_text(size=30))+
ggtitle("BWM")
dev.off()
```

##Plot hyp tm D1 vs 11, AL vs DR#\#

```{r}
data_aldr_youngold_hypNAremov <- data_aldr_youngold[data_aldr_youngold$t.h.tm != "#N/A", ] 
data_aldr_youngold_hypNAremov$t.h.tm <- as.numeric(data_aldr_youngold_hypNAremov$t.h.tm)

pdf(paste(Sys.Date(),"AL DR Y O hyp.pdf"))
ggplot(data=data_aldr_youngold_hypNAremov, aes(x=Age, y=t.h.tm, fill=Nutrients))+
geom_violin()+
stat_summary(fun.data="mean_cl_boot", size=1, position = position_dodge(width = 0.9))+
ylab(bquote("B.A./NAD(P)H  " *tau[m]* " (ns)"))+
xlab("Age (Days)")+
theme_classic(base_size = 20, base_family = "")+
theme(aspect.ratio=1)+
theme(text=element_text(size=30))+
ggtitle("Hypodermis")
dev.off()
```

##Plot pm tm D1 vs 11, AL vs DR#\#

```{r}
data_aldr_youngold_pmNAremov <- data_aldr_youngold[data_aldr_youngold$pm.tm != "#N/A", ] 
data_aldr_youngold_pmNAremov$pm.tm <- as.numeric(data_aldr_youngold_pmNAremov$pm.tm)

pdf(paste(Sys.Date(),"AL DR Y O pm.pdf"))
ggplot(data=data_aldr_youngold_pmNAremov, aes(x=Age, y=pm.tm, fill=Nutrients))+
geom_violin()+
stat_summary(fun.data="mean_cl_boot", size=1, position = position_dodge(width = 0.9))+
ylab(bquote("B.A./NAD(P)H  " *tau[m]* " (ns)"))+
xlab("Age (Days)")+
theme_classic(base_size = 20, base_family = "")+
theme(aspect.ratio=1)+
theme(text=element_text(size=30))+
ggtitle("PM")
dev.off()
```

##Plot germ tm D1 vs 9, AL vs DR#\#
```{r}
myagesoygerm <- c("1","9")
data_aldr_youngoldgerm <- as.data.frame(subset(data_mito_animal_matrix, Age%in%myagesoygerm))
data_aldr_youngoldgerm$Age <- as.character(data_aldr_youngoldgerm$Age)

data_aldr_youngoldgerm_germNAremov <- data_aldr_youngoldgerm[data_aldr_youngoldgerm$g.tm != "#N/A", ] 
data_aldr_youngoldgerm_germNAremov$g.tm <- as.numeric(data_aldr_youngoldgerm_germNAremov$g.tm)

pdf(paste(Sys.Date(),"AL DR Y O germ.pdf"))
ggplot(data=data_aldr_youngoldgerm_germNAremov, aes(x=Age, y=g.tm, fill=Nutrients))+
geom_violin()+
stat_summary(fun.data="mean_cl_boot", size=1, position = position_dodge(width = 0.9))+
ylab(bquote("B.A./NAD(P)H  " *tau[m]* " (ns)"))+
xlab("Age (Days)")+
theme_classic(base_size = 20, base_family = "")+
theme(aspect.ratio=1)+
theme(text=element_text(size=30))+
ggtitle("Germline")
dev.off()
```

##PCA of worm body wall mucsle Y O AL DR#\#
```{r}
set.seed(1)
myvars_mus <- c("t.bwm.a1","t.bwm.t1","t.bwm.t2","t.bwm.i","Condition")
data_aldr_youngold_mus_pca <- data_aldr_youngold[myvars_mus]
data_aldr_youngold_mus_pca_NAremov <- data_aldr_youngold_mus_pca[data_aldr_youngold_mus_pca$t.bwm.a1 != "#N/A",]

data_aldr_youngold_mus_pca_NAremov$t.bwm.a1s <- scale(as.numeric(data_aldr_youngold_mus_pca_NAremov$t.bwm.a1))
data_aldr_youngold_mus_pca_NAremov$t.bwm.t1s <- scale(as.numeric(data_aldr_youngold_mus_pca_NAremov$t.bwm.t1))
data_aldr_youngold_mus_pca_NAremov$t.bwm.t2s <- scale(as.numeric(data_aldr_youngold_mus_pca_NAremov$t.bwm.t2))
data_aldr_youngold_mus_pca_NAremov$t.bwm.is <- scale(as.numeric(data_aldr_youngold_mus_pca_NAremov$t.bwm.i))

pca <- prcomp(data_aldr_youngold_mus_pca_NAremov[,6:9], scale=FALSE)
pcadatamus <- data.frame(pca$x[,1], pca$x[,2], Condition=data_aldr_youngold_mus_pca_NAremov$Condition)

pdf(paste(Sys.Date(),"bwm pca al dr y o.pdf"))
ggplot(pcadatamus, aes(x=pca$x[,1], y=pca$x[,2], col=Condition))+
  geom_point(size = 2)+
  theme_classic(base_size = 25, base_family = "")+
  ylab(bquote("PC2"))+
  xlab(bquote("PC1"))+
  theme(aspect.ratio = 1)+
  ggtitle("BWM")#+
dev.off()
```

##PCA of worm pm Y O AL DR#
```{r}
set.seed(1)
myvars_pm <- c("pm.a1","pm.t1","pm.t2","pm.i","Condition")
data_aldr_youngold_pm_pca <- data_aldr_youngold[myvars_pm]
data_aldr_youngold_pm_pca_NAremov <- data_aldr_youngold_pm_pca[data_aldr_youngold_pm_pca$pm.a1 != "#N/A",]

data_aldr_youngold_pm_pca_NAremov$pm.a1s <- scale(as.numeric(data_aldr_youngold_pm_pca_NAremov$pm.a1))
data_aldr_youngold_pm_pca_NAremov$pm.t1s <- scale(as.numeric(data_aldr_youngold_pm_pca_NAremov$pm.t1))
data_aldr_youngold_pm_pca_NAremov$pm.t2s <- scale(as.numeric(data_aldr_youngold_pm_pca_NAremov$pm.t2))
data_aldr_youngold_pm_pca_NAremov$pm.is <- scale(as.numeric(data_aldr_youngold_pm_pca_NAremov$pm.i))

pca <- prcomp(data_aldr_youngold_pm_pca_NAremov[,6:9], scale=FALSE)
pcadatapm <- data.frame(pca$x[,1], pca$x[,2], Condition=data_aldr_youngold_pm_pca_NAremov$Condition)

pdf(paste(Sys.Date(),"pm pca al dr y o.pdf"))
ggplot(pcadatapm, aes(x=pca$x[,1], y=pca$x[,2], col=Condition))+
  geom_point(size = 2)+
  theme_classic(base_size = 25, base_family = "")+
  ylab(bquote("PC2"))+
  xlab(bquote("PC1"))+
  theme(aspect.ratio = 1)+
  ggtitle("PM")
dev.off()
```

##PCA of worm hyp Y O AL DR
```{r}
set.seed(1)
myvars_hyp <- c("t.h.a1","t.h.t1","t.h.t2","t.h.i","Condition")
data_aldr_youngold_hyp_pca <- data_aldr_youngold[myvars_hyp]
data_aldr_youngold_hyp_pca_NAremov <- data_aldr_youngold_hyp_pca[data_aldr_youngold_hyp_pca$t.h.a1 != "#N/A",]

data_aldr_youngold_hyp_pca_NAremov$t.h.a1s <- scale(as.numeric(data_aldr_youngold_hyp_pca_NAremov$t.h.a1))
data_aldr_youngold_hyp_pca_NAremov$t.h.t1s <- scale(as.numeric(data_aldr_youngold_hyp_pca_NAremov$t.h.t1))
data_aldr_youngold_hyp_pca_NAremov$t.h.t2s <- scale(as.numeric(data_aldr_youngold_hyp_pca_NAremov$t.h.t2))
data_aldr_youngold_hyp_pca_NAremov$t.h.is <- scale(as.numeric(data_aldr_youngold_hyp_pca_NAremov$t.h.i))

pca <- prcomp(data_aldr_youngold_hyp_pca_NAremov[,6:9], scale=FALSE)
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)

pcadatahyp <- data.frame(pca$x[,1], pca$x[,2], Condition=data_aldr_youngold_hyp_pca_NAremov$Condition)

pdf(paste(Sys.Date(),"hyp pca al dr y o.pdf"))
ggplot(pcadatahyp, aes(x=pca$x[,1], y=pca$x[,2], col=Condition))+
  geom_point(size = 2)+
  theme_classic(base_size = 25, base_family = "")+
  ylab(bquote("PC2"))+
  xlab(bquote("PC1"))+
  theme(aspect.ratio = 1)+
  ggtitle("Hypodermis")
dev.off()
```

##Plot tm line AL vs DR across age for muscle#\#
```{r}
mytissuesmus <- c("Caudal BWM")
datamus <- as.data.frame(subset(data_averages, TissueFullName%in%mytissuesmus))

pdf(paste(Sys.Date(),"bwm tm line.pdf"))
ggplot(data=datamus, aes(x=Age, y=tmmean, color=Nutrients))+
geom_line(size=2)+
theme_classic(base_size = 20, base_family = "")+
theme(legend.position = "none")+
ylab(bquote("B.A./NAD(P)H  " *tau[m]* " (ns)"))+
xlab("Age (Days)")+
theme(aspect.ratio=1)+
theme(text=element_text(size=25))+ 
theme(legend.position="right")+
ggtitle("BWM")
dev.off()
```

##Plot tm line AL vs DR across age for hypodermis#\#

```{r}
mytissueshyp <- c("Caudal Hypodermis")
datahyp <- as.data.frame(subset(data_averages, TissueFullName%in%mytissueshyp))

pdf(paste(Sys.Date(),"hyp tm line.pdf"))
ggplot(data=datahyp, aes(x=Age, y=tmmean, color=Nutrients))+
geom_line(size=2)+
theme_classic(base_size = 20, base_family = "")+
theme(legend.position = "none")+
ylab(bquote("B.A./NAD(P)H  " *tau[m]* " (ns)"))+
xlab("Age (Days)")+
theme(aspect.ratio=1)+
theme(text=element_text(size=25))+ 
theme(legend.position="right")+
ggtitle("Hypodermis")
dev.off()
```

##Plot tm line AL vs DR across age for Pharyngeal Muscle#\#

```{r}
mytissuespm <- c("PM")
datapm <- as.data.frame(subset(data_averages, TissueFullName%in%mytissuespm))

pdf(paste(Sys.Date(),"pm tm line.pdf"))
ggplot(data=datapm, aes(x=Age, y=tmmean, color=Nutrients))+
geom_line(size=2)+
theme_classic(base_size = 20, base_family = "")+
theme(legend.position = "none")+
ylab(bquote("B.A./NAD(P)H  " *tau[m]* " (ns)"))+
xlab("Age (Days)")+
theme(aspect.ratio=1)+
theme(text=element_text(size=25))+ 
theme(legend.position="right")+
ggtitle("PM")
dev.off()
```

##Plot tm line AL vs DR across age for Germline#\#

```{r}
mytissuesgerm <- c("Germline")
myages_no11 <- c("1","3","5","7","9")

datagerm <- as.data.frame(subset(data_averages, TissueFullName%in%mytissuesgerm))
datagerm <- as.data.frame(subset(datagerm, Age%in%myages_no11))

pdf(paste(Sys.Date(),"germ tm line.pdf"))
ggplot(data=datagerm, aes(x=Age, y=tmmean, color=Nutrients))+
geom_line(size=2)+
theme_classic(base_size = 20, base_family = "")+
theme(legend.position = "none")+
ylab(bquote("B.A./NAD(P)H  " *tau[m]* " (ns)"))+
xlab("Age (Days)")+
theme(aspect.ratio=1)+
theme(text=element_text(size=25))+ 
theme(legend.position="right")+
ggtitle("Germline")
dev.off()
```
